Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-29 soc_squeezing/cumulant_expansions.tex
add more thorough explanation of decoherence calculation method

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r,
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonians
for one-axis twisting with a transverse field, for example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_z^2 + \Omega S_x
  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{z,y}
  = \chi \p{S_z^2 - S_y^2}
  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{y,x}
  = \chi\p{S_y^2 - S_x^2}
  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r.
\end{align}
Equipped with a matrix representation of the time derivative operator,
i.e.~$T^{\ell m n}_{pqr}$, we can expand
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
  \label{eq:time_series}
\end{align}
where
\begin{align}
  \sp{T^k}^{\ell mn}_{pqr}
  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
\end{align}
As terms in \eqref{eq:time_series} are suppressed factorially in $k$,
for any fixed (or maximal) time $t$ we can truncate the series to
write
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  \approx \sum_{k=0}^{M_t-1} \f{t^k}{k!}
  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
\end{align}
where $M_t$ is chosen such that
\begin{align}
  \f{t^{M_t}}{M_t!} \sum_{p,q,r}
  \sp{T^{M_t}}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0}
  \ll S^{\ell+m+n}.
\end{align}
The commutator and decoherence operator necessary for computing the
time derivative operator $T^{\ell m n}_{pqr}$ are respectively worked
out in Appendix \ref{sec:comm_general} and \ref{sec:decoherence},
while expectation values for initial conditions of interest are
provided in Appendix \ref{sec:initial_conditions}.
% comment on explosion of operators

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_mu_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively determine commutation
relations with higher-weight spin operators.  For example, we can find
that
\begin{align}
  S_z S_\mu^m
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
  = S_\mu^m \p{\mu m + S_z}.
  \label{eq:push_z_mu_single}
\end{align}
Repeating this process $\ell$ times tells us
\begin{align}
  S_z^\ell S_\mu^m = S_\mu^m \p{\mu m + S_z}^\ell,
  \label{eq:push_z_mu}
\end{align}
which implies
\begin{align}
  \sp{S_z^\ell, S_\mu^m}_-
  = S_\mu^m \sp{\p{\mu m + S_z}^\ell - S_z^\ell}
  = S_\mu^m \sum_{k=0}^{\ell-1} {\ell \choose k} \p{\mu m}^{\ell-k} S_z^k.
  \label{eq:comm_z_mu}
\end{align}
Similarly finding commutation relations between high-weight transverse
spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
considerably more difficult than the cases we have worked out thus
far.  We therefore save this work for Appendix
\ref{sec:comm_transverse}.

@@

To determine commutation relations between high-weight collective spin
operators transverse to the decoherence (i.e.~$z$) axis, we compute
(with $\bmu\equiv-\mu\in\set{+1,-1}$)

@@

  S_\mu S_\bmu^n
  &= S_\bmu S_\mu S_\bmu^{n-1} + 2\mu S_z S_\bmu^{n-1} \\
  &= S_\bmu \sp{S_\bmu S_\mu S_\bmu^{n-2} + 2\mu S_z S_\bmu^{n-2}}
  + 2\mu S_z S_\bmu^{n-1} \\
  &= S_\bmu^2 S_\mu S_\bmu^{n-2}
  + 2\mu \p{S_\bmu S_z S_\bmu ^{n-2} + S_z S_\bmu^{n-1}} \\
  &= S_\bmu^n S_\mu + 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}.
  \label{eq:comm_mu_single}

@@

While \eqref{eq:comm_mu_single} gives us the commutator
$\sp{S_\mu,S_\bmu^n}_-$, we would like to enforce an ordering on
products of collective spin operators, which allows us to only keep
track of high-weight operators that are linearly independent.  We
choose to impose an ordering for which all collective spin operators
take the form $S_\bmu^\ell S_\mu^m S_z^n$, as such an ordering will
prove convenient for the calculations in this section.  This choice of

@@

  2 \mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
  = 2 \mu \sum_{k=0}^{n-1}
  S_\bmu^k S_\bmu^{n-k-1} \p{\bmu\sp{n-k-1} + S_z}
  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z},

@@

so

@@

  \sp{S_\mu, S_\bmu^n}_-
  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}.

@@

together with \eqref{eq:C_mn_E} implies that

@@

\section{Decoherence of collective spin operators through individual
  spin decay}
\label{sec:decoherence}

@@

If we consider only symmetrically ordered collective spin operators,
there are two cases to work out: the effect of decoherence through
$\D_\mu$ on $S_\mu^\ell S_z^m S_\bmu^n$, and the effect on
$S_\bmu^\ell S_z^m S_\mu^n$.  We work out the latter case first,
namely
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
  = \sum_j\sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
  - S S_\bmu^\ell S_z^m S_\mu^n
  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+.
\end{align}
To simplify the first term, we use \eqref{eq:comm_z_mu_base} to push
the jump operators together, finding
\begin{align}
  \sum_j \sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
  = S_\bmu^\ell \sum_j \sigma_j^\bmu \sigma_j^\mu \p{\mu+S_z}^m S_\mu^n
  = S_\bmu^\ell \p{S+S_z} \p{\mu+S_z}^m S_\mu^n.
\end{align}
We can likewise push $S_z$ operators together using
\eqref{eq:push_z_mu_single} to expand
\begin{align}
  \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
  = S_z S_\bmu^\ell S_z^m S_\mu^n + S_\bmu^\ell S_z^m S_\mu^n S_z
  = S_\bmu^\ell\p{\bmu\ell - \mu n + 2 S_z} S_z^m S_\mu^n,
\end{align}
which implies
\begin{align}
  \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
  = S_\bmu^\ell\p{-\f12\sp{\ell+n} + \mu S_z} S_z^m S_\mu^n,
\end{align}
and in turn
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
  &= S_\bmu^\ell \sp{\p{S+S_z} \p{\mu+S_z}^m
    - \p{S + \f12\sp{\ell+n} + S_z} S_z^m} S_\mu^n \\
  &= S_\bmu^\ell
  \sp{\p{S+S_z} \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k} S_z^k
  - \f12\p{\ell+n} S_z^m} S_\mu^n.
\end{align}
Decoherence through $\D_\mu$ only couples operators of the form
$S_\bmu^\ell S_z^m S_\mu^n$ to operators of the (almost identical)
form $S_\bmu^\ell S_z^{m'} S_\mu^n$ with $m'\le m$.

@@

Initial conditions for all collective spin operators can be determined
from the expansions
\begin{align}
  S_z^n = \sum_{k=-S}^S k^n \op{k},
  &&
  S_\mu
  = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{k,1}}
  g_\mu\p{k} \op{k+\mu}{k},
\end{align}
with
\begin{align}
  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
\end{align}
and the initial states
\begin{align}
  \ket{-Z} \equiv \ket{-S},
  &&
  \ket{X} \equiv 2^{-N/2} \sum_{k=-S}^S { N \choose S+k }^{1/2} \ket{k},
\end{align}
using which we can compute
\begin{align}
  \bk{-Z|S_+^\ell S_z^m S_-^n|-Z}
  = \delta_{\ell,0} \delta_{m,0} \p{-S}^m,
\end{align}
and
\begin{align}
  \bk{X|S_+^\ell S_z^m S_-^n|X}
  &=  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \tag*{} \\
  &\qquad\qquad\qquad\qquad
  \times \prod_{p=0}^{\ell-1} g_+\p{S,k+p}
  \prod_{q=0}^{n-1} g_-\p{S,k+n-q} \\
  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \f{\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
\end{align}
With the help of Mathematica, we can evaluate the sum to find
\begin{align}
  \bk{X|S_+^\ell S_z^m S_-^n|X}
  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
\end{align}
where $\alpha$ and $\beta$ are defined in terms of the generalized
hypergeometric function $\F$ by
\begin{align}
  \alpha
  \equiv \F\p{\let\scriptstyle\textstyle
    \substack{\ell-N,~n-N\\-N}~; -1},
\end{align}
and
\begin{align}
  \beta
  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
    \substack{1+\ell-N,~1+n-N\\1-N}~; -1},
\end{align}
for
\begin{align}
  \tilde\F_q\p{\let\scriptstyle\textstyle
    \substack{a,~b\\c}~; z}
  \equiv \F\p{\let\scriptstyle\textstyle
    \substack{2,2,\cdots,2,a,b\\1,1,\cdots,1,c}~; z}
\end{align}
with $q$ repetitions of 2 and 1.

>>>>

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r,
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonians
for one-axis twisting with a transverse field, for example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_z^2 + \Omega S_x
  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{z,y}
  = \chi \p{S_z^2 - S_y^2}
  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{y,x}
  = \chi\p{S_y^2 - S_x^2}
  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
  \label{eq:time_derivative}
\end{align}
where commutator and decoherence operator necessary for computing the
matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
$d/dt$ are respectively worked out in Appendix \ref{sec:comm_general}
and \ref{sec:decoherence}.

The time derivative operator $d/dt$ will generally couple spin
operators $S_+^\ell S_z^m S_-^n$ to spin operators $S_+^p S_z^q S_-^r$
with higher weight, i.e.~with $p+q+r>\ell+m+n$.  In practice, keeping
track of this operator growth will eventually require more
computational resources than are available, so we must truncate it at
some point by taking
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
  \label{eq:time_derivative_truncation}
\end{align}
for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$,
although the most optimal weight function will generally be
system-dependent.  The truncation in
\eqref{eq:time_derivative_truncation} will close the system of
differential equations defined by \eqref{eq:time_derivative}, and
allow us to solve it using standard numerical methods.  The
expectation values of collective spin operators for initial conditions
of interest are provided in Appendix \ref{sec:initial_conditions}.

Alternately, we can formally solve the differential equation in
\eqref{eq:time_derivative} and write
\begin{align}
  S_+^\ell S_z^m S_-^n
  = \left.\exp\p{t \f{d}{dt}}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n},
\end{align}
which implies
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements of the $k$-th time derivative are given by
\begin{align}
  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
  &&
  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
\end{align}
\begin{align}
  T^{\ell mn}_{pqr;k>1}
  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
\end{align}
Instead of explicitly choosing a weight function that determines which
operators to keep track of, we can now truncate the series in
\eqref{eq:time_series} by taking
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}
  \to \sum_{k<M} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}.
\end{align}
Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
$k=1,2,\cdots,M-1$ now {\it tells} us which operators we need to keep
track of in order to compute the expectation value
$\bk{S_+^\ell S_z^m S_-^n}$.  Furthermore, after computing the
expectation values of all relevant operators at $t=0$, time evolution
of $\bk{S_+^\ell S_z^m S_-^n}$ takes only $\O\p{M}$ CPU time.

As a final comment, we note that the relation
$\p{S_+^\ell S_z^m S_-^n}^\dag=S_+^n S_z^m S_-^\ell$ can be exploited
to cut the number of operators we need to keep track of roughly in
two.

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively compute identities
involving powers of collective spin operators.  Working through the
possibilities, we find
\begin{align}
  S_z^m \sigma_j^\mu
  = S_z^{m-1} \sigma_j^\mu \p{\mu + S_z}
  = S_z^{m-2} \sigma_j^\mu \p{\mu + S_z}^2
  = \cdots
  = \sigma_j^\mu \p{\mu + S_z}^m,
  \label{eq:push_z_mu_Ss}
\end{align}
and
\begin{align}
  \sigma_j^z S_\mu^m
  = \p{2\mu\sigma_j^\mu + S_\mu \sigma_j^z} S_\mu^{m-1}
  = 2\mu\sigma_j^\mu S_\mu^{m-1} + S_\mu \sigma_j^z S_\mu^{m-1}
  = \cdots
  = 2\mu m\sigma_j^\mu S_\mu^{m-1} + S_\mu^m \sigma_j^z.
  \label{eq:push_z_mu_sS}
\end{align}
Summing over the single-spin index $j$ in both of these cases gives us
the purely collective-spin versions of these identities:
\begin{align}
  S_z^m S_\mu = S_\mu \p{\mu + S_z}^m,
  &&
  S_z S_\mu^m = S_\mu^m \p{\mu m + S_z},
  \label{eq:push_z_mu_single}
\end{align}
where we can repeat the process of pushing through individual
operators $\ell$ times to get
\begin{align}
  S_z^\ell S_\mu^m
  = S_z^{\ell-1} S_\mu^m \p{\mu m + S_z}
  = S_z^{\ell-2} S_\mu^m \p{\mu m + S_z}^2
  = \cdots
  = S_\mu^m \p{\mu m + S_z}^\ell.
  \label{eq:push_z_mu}
\end{align}
Similarly finding commutation relations between powers of transverse
spin operators, i.e.~$S_\mu$, turns out to be considerably more
difficult than the cases we have worked out thus far.  We therefore
save this work for Appendix \ref{sec:comm_transverse}.

@@

To find commutation relations between powers of transverse collective
spin operators, we compute (with $\bmu\equiv-\mu\in\set{+1,-1}$)

@@

  S_\mu^m \sigma_j^\bmu
  &= S_\mu^{m-1}\sigma_j^\bmu S_\mu + \mu S_\mu^{m-1} \sigma_j^z \\
  &= S_\mu^{m-2} \sigma_j^\bmu S_\mu^2 + \mu S_\mu^{m-2} \sigma_j^z S_\mu
  + \mu S_\mu^{m-1} \sigma_j^z \\
  &= \sigma_j^\bmu S_\mu^m
  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
  \label{eq:push_mu_Ss_start}.

@@

While \eqref{eq:push_mu_Ss_start} gives us the commutator
$\sp{S_\mu^m,\sigma_j^\bmu}_-$, we would like to enforce an ordering
on products of spin operators, which will allow us to only keep track
of operators that are linearly independent.  We choose to impose an
ordering with all $\sigma_j^\bmu$ operators on the left, and all
$\sigma_j^z$ operators on the right, as such an ordering will prove
convenient for the calculations in this section.  This choice of

@@

  \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
  &= \mu \sum_{k=0}^{mq-1} S_\mu^k
  \p{2\mu\sp{m-k-1} \sigma_j^\mu S_\mu^{m-k-2}
    + S_\mu^{m-k-1} \sigma_j^z} \\
  &= m \p{m-1} \sigma_j^\mu S_\mu^{m-2} + \mu m S_\mu^{m-1} \sigma_j^z,

@@

@@

  S_\mu^m \sigma_j^\bmu
  = \sigma_j^\bmu S_\mu^m + m \p{m-1} \sigma_j^\mu S_\mu^{m-2}
  + \mu m S_\mu^{m-1} \sigma_j^z,
  \label{eq:push_mu_Ss}
and in turn
  S_\mu^m S_\bmu = S_\bmu S_\mu^m
  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_z}.
  \label{eq:push_mu_single}

@@

together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
implies that

@@

\section{Decoherence of collective spin operators from uncorrelated
  spin decay}
\label{sec:decoherence}

@@

If we consider only symmetrically ordered collective spin operators,
there are two cases to work out: the effect of decoherence through
$\D_\mu$ on $S_\mu^\ell S_z^m S_\bmu^n$, and the effect on
$S_\bmu^\ell S_z^m S_\mu^n$.  We work out the latter case first,
namely
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
  = \sum_j\sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
  - S S_\bmu^\ell S_z^m S_\mu^n
  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+.
\end{align}
To simplify the first term, we use \eqref{eq:comm_z_base} to push the
jump operators together, finding
\begin{align}
  \sum_j \sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
  = S_\bmu^\ell \sum_j \sigma_j^\bmu \sigma_j^\mu \p{\mu+S_z}^m S_\mu^n
  = S_\bmu^\ell \p{S+S_z} \p{\mu+S_z}^m S_\mu^n.
\end{align}
We can likewise push $S_z$ operators together using
\eqref{eq:push_z_mu_single} to expand
\begin{align}
  \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
  = S_z S_\bmu^\ell S_z^m S_\mu^n + S_\bmu^\ell S_z^m S_\mu^n S_z
  = S_\bmu^\ell\p{\bmu\ell - \mu n + 2 S_z} S_z^m S_\mu^n,
\end{align}
which implies
\begin{align}
  \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
  = S_\bmu^\ell\p{-\f12\sp{\ell+n} + \mu S_z} S_z^m S_\mu^n,
\end{align}
and in turn
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
  &= S_\bmu^\ell \sp{\p{S+S_z} \p{\mu+S_z}^m
    - \p{S + \f12\sp{\ell+n} + S_z} S_z^m} S_\mu^n \\
  &= S_\bmu^\ell
  \sp{\p{S+S_z} \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k} S_z^k
  - \f12\p{\ell+n} S_z^m} S_\mu^n.
\end{align}
Decoherence through $\D_\mu$ only couples operators of the form
$S_\bmu^\ell S_z^m S_\mu^n$ to operators of the (almost identical)
form $S_\bmu^\ell S_z^{m'} S_\mu^n$ with $m'\le m$.

@@

Initial conditions for all collective spin operators can be determined
from the expansions
\begin{align}
  S_z^n = \sum_{k=-S}^S k^n \op{k},
  &&
  S_\mu
  = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{k,1}}
  g_\mu\p{k} \op{k+\mu}{k},
\end{align}
with
\begin{align}
  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
\end{align}
and the initial states
\begin{align}
  \ket{\pm Z} \equiv \ket{\pm S},
  &&
  \ket{X} \equiv 2^{-N/2} \sum_{k=-S}^S { N \choose S+k }^{1/2} \ket{k},
\end{align}
using which we can compute
\begin{align}
  \bk{\pm Z|S_+^\ell S_z^m S_-^n|\pm Z}
  = \delta_{\ell,0} \delta_{m,0} \p{\pm S}^m,
\end{align}
and
\begin{align}
  \bk{X|S_+^\ell S_z^m S_-^n|X}
  &=  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \tag*{} \\
  &\qquad\qquad\qquad\qquad
  \times \prod_{p=0}^{\ell-1} g_+\p{S,k+p}
  \prod_{q=0}^{n-1} g_-\p{S,k+n-q} \\
  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \f{\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
\end{align}
With the help of Mathematica, we can write the sum in the alternate
form
\begin{align}
  \bk{X|S_+^\ell S_z^m S_-^n|X}
  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
\end{align}
where $\alpha$ and $\beta$ are defined in terms of the generalized
hypergeometric function $\F$ by
\begin{align}
  \alpha
  \equiv \F\p{\let\scriptstyle\textstyle
    \substack{\ell-N,~n-N\\-N}~; -1},
\end{align}
and
\begin{align}
  \beta
  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
    \substack{1+\ell-N,~1+n-N\\1-N}~; -1},
\end{align}
for
\begin{align}
  \tilde\F_q\p{\let\scriptstyle\textstyle
    \substack{a,~b\\c}~; z}
  \equiv \F\p{\let\scriptstyle\textstyle
    \substack{2,2,\cdots,2,a,b\\1,1,\cdots,1,c}~; z}
\end{align}
with $q$ repetitions of 2 and 1.
